Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-09-21 paper.tex
completed toil summary

\newcommand{\note}[1]{\textcolor{green}{\textit{note}: #1}}
\newcommand{\tristan}[1]{\textcolor{red}{TG: #1}}
\newcommand{\weird}[1]{\uwave{#1}}

@@

    \section{Geneflow}
        \subsection{Pipeline description}
        \subsection{Pipeline execution}
        \subsection{Debugging and provenance}
    \section{Toil}
        \subsection{Pipeline description}
        \subsection{Pipeline execution}
        \subsection{Debugging and provenance}
    \section{Thunder}

>>>>

\newcommand{\note}[1]{\textcolor{blue}{\textit{note}: #1}}
\newcommand{\tristan}[1]{\textcolor{red}{TG: #1}}
\newcommand{\weird}[1]{\uwave{#1}}

@@

        \section{Nextflow} \subsection{summary} Nextflow is a POSIX-compatible
        pipeline engine for scientific workflows.  It extends the Groovy
        programming language to provide an  intuitive domain-specific language
        (DSL) that enables users to easily describe their workflows. Unlike
        pipeline engines such as Nipype and PSOM, Netflow does not employ DAG.
        It instead adheres to the \textit{Dataflow} programming model which
        decides how data will flow through the pipeline at runtime \note{fix
        this sentence}. Nextflow prides itself on allowing the integration of
        containerized tools to address the issue of numerical instability that
        may arise from different environment setups.

    \section{Geneflow} 
        \subsection{Pipeline description} 
        \subsection{Pipeline execution} 
        \subsection{Debugging and provenance} 
    \section{Toil}
        \note{obtained from paper's supplementary documentation}
        Toil is another general workflow framework aimed at the processing of 
        scientific big data, with its main focus being on genomics workflows. It 
        prides itself in not only its ability to not only scale the processing 
        of very large datasets, but also its portability, allowing it to be 
        executed on any environment. Toil pipelines can be written in CWL, WDL 
        (Workflow Description Language) and Python. The Toil workflow 
        framework is also open-sourced, making it a community-driven effort. 
        \note{as is the case with most}

        \subsection{Pipeline description}
           A Toil workflow consists of a DAG of Jobs connected to each other
           through input and output dependencies. All custom user scripts must
           inherit from the Job class. In order to absolve the user from having
           to manually create these classes for each function, Toil provides
           wrapper classes that convert Python functions to Toil jobs directly.
           A Toil job need not be part of the static workflow, it can also be
           added at runtime.

           There exists three types of relationships between Toil jobs: parent, 
           child and follow-on jobs. Child jobs begin execution immediately
           after their Parent completes their execution. Follow-on jobs, however
           , begin execution after the Parent, associated child job and their
           successors complete execution. They are particularly useful as 
           cleanup jobs. Both child jobs and follow-on jobs are executed in
           parallel.

           Variables may be passed at runtime to dynamically added child or
           follow-on jobs, or they can be referenced statically using promises.
           Promises enable the reference to the result to be passed prior to 
           runtime, evaluating the result only at runtime.

           Workflows can be executed or 
           resumed through the Runner class, or alternatively, invoked by the
           Toil class. The latter differing from the Runner class as it serves 
           as the workflow's context manager. The Toil class performs
           a preliminary configuration of the workflow (e.g. file staging).
           All Toil workflows begin with a single job. The value of this root 
           job is returned at the end of the workflow.

            The FileStore class manages the permanent and temporary files
            created by the workflow. This class ensures that all necessary files
            will be cleaned up after job execution, but also that workflows
            can resume from the last successful state. Both temporary 
            directories and temporary files can be created at a user's desired
            location (e.g. local file system, Amazon S3, etc.,). Should files
            and directories need to persist during workflow execution and
            be accessible to all jobs, the FileStore
            can add these files to a global store. All files stored in the 
            global store are immutable, preventing any updates to these files
            from occuring other than deletion. There also does not exist any
            kind of filesystem hierarchy in the global store.\note{allows use
            of various different object stores/ enables caching of files as they
            never change}

            In order to add files to a workflow, it is necessary to stage them
            prior to execution. This is achieved using the \textit{Toil} context
            manager as it is running on the leader prior to execution. It is 
            possible to stage the file as a shared file, if necessary. In the 
            instance of a job failure, the staged file behaves like any other 
            file in the job store. If the file is not cleaned up post execution,
            the file will not need to be staged again during resumption.

            Toil also provides the functionality to startup services, such as
            databases or servers, during pipeline execution. Multiple services
            can be define per job and services can have associated child
            services. This functionality allows, for instance, Apache Spark 
            clusters to be spawned during workflow execution. As services 
            complicate workflow resumption post failure, Toil enables the
            checkpointing of services such that the workflow will resume from a
            checkpoint if any of the successors to the service fail to restart.

        \subsection{Pipeline execution} 
            \note{obtained from website documentation}
            Toil is composed of three components: 1) the Job Store, 2) the Batch
            System and 3) the Provisioner. The Job Store centralizes all the 
            files necessary for the workflow. During a workflow re-execution,
            the job store is used to resume the workflow from its last
            successful task. Currently, there are two types of job stores 
            supported: the file job store and the cloud job store.

            The Batch System specifies which type of scheduler will be used.
            A Toil workflow can either be executed locally, on HPC clusters or 
            the cloud.

            The Provisioner, on the other hand, is used to execute Toil
            workflows on specific cloud platforms. Using the Cluster Utilities
            command line tools, a user can provision nodes for their desired 
            cloud platform.

            Toil executes workflows using a leader-worker architecture. The
            leader is a single-threaded read-only process responsible for 
            determining which jobs are ready to be executed. Workers, on the 
            other hand, are created by the batch-system and are responsible
            for the execution of a job. To alleviate the single-threaded leader
            from having to do an excessive amount of work, the worker maintains
            track of its status (i.e. success or failure) and reports its status
            back to the leader by writing it to the job-store. The worker may
            also choose to immediately execute any child-jobs defined by the 
            job it has executed.

            In the instance when a job or a subset of jobs fail, the workflow
            will continue to execute until it can proceed no more and will throw
            an error. The user is given the option to resolve the bug and resume
            the workflow or delete the job store and reexecute the pipeline 
            entirely.
        \subsection{Debugging and provenance}

        By default, workflow standard input and output is hidden to the user.
        It is however possible to make the log data visible to the user during
        execution or to redirect this data to a logfile.

        Toil also provide a fuctionality, stats, to capture statistics on the 
        workflow in the job store. Examples of the statistics it captures
        includes cpu, memory and job duration.

    \section{Thunder}
